# Assessment of exercise activity using the data of activity monitors
Aleksandra Lyulko

November 22, 2013

## Summary

The goal of this study is to predict whether a subject makes one of the common
mistakes while lifting a dumbell. The mistakes and the correct exectuion are
summarized in a variable "classe" of the training dataset. I used Prediction
with Classification Trees ('rpart' package) and the Linear Discriminant
Analysis (LDA) to build classification models on the training data and 
evaluated their accuracies.
I also tried the Linear Model - it turned out to be not suitable for factor
outcomes; Random Forest, Naive Bayes and Gradient Boosting Machine - for these
methods the computation time was too long.
Based on the accuracies (50% for Trees and 62% for LDA) I chose LDA to apply
to the 20 test sets.

## Data Processing

### Download the data

```{r loading, cache=TRUE}
url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url1="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
setInternet2(use=TRUE)
download.file(url, destfile="pml-training.csv")
download.file(url1, destfile="pml-testing.csv")

training=read.csv("pml-training.csv")
testing=read.csv("pml-testing.csv")
```
### Select predictors
Select available variables for training, based on the available values in the
test set and exclude the variables related to the observation number, subject
name and time:

```{r cleaning, cache=TRUE}
nzv = nearZeroVar(testing, saveMetrics=TRUE)
nzv1 = nearZeroVar(training, saveMetrics=TRUE)
indices = which(nzv$zeroVar=="FALSE")
training1 = training[,indices]
testing1 = testing[,indices]
testing2 = subset(testing1, select=-c(X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, problem_id))
training2 = subset(training1, select=-c(X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))
```

### Building the models
Training using the Classification with Trees model:
```{r Treetraining, cache=TRUE}
modFitTree = train(classe~., method="rpart", data=training2)
predTree = predict(modFitTree, training2)
```

Training with LDA:
```{r Ldatraining, cache=TRUE}
modFitLda=train(classe~., data=training2, method="lda")
predLda=predict(modFitLda, training2)
```

### Cross-validation
Validate the models with the number of correctly predicted training cases:
```{r Nvalidation}
TreeValid=(predTree == training2$classe)
table(TreeValid)
LdaValid = (predLda == training2$classe)
table(LdaValid)
```
This simple test shows that performance of the LDA model is significantly
better than Tree classification.

Validate the models using the Confusion Matrix:
```{r CMvalidation}
confusionMatrix(predTree, training2$classe)
confusionMatrix(predLda, training2$classe)
```
The accuracy of the Classification Tree model is 0.4956 with sensitivities
for different classes ranging from 0.0 to 0.9104 (0.9104 for class A). The
accuracy of the LDA model is 0.7161 with sensitivities from 0.6310 to 0.8267.
Thus, I decided to use the Linear Discriminant Analysis as my model to predict
the outcomes of the test sets.

I estimate the out-of-sample error to be below 72%, based on the Accuracy of
LDA model.
 

## Results
Apply the LDA model to the testing set to predict the manner of
exercise of the participants in the test cases.
```{r predict}
predLdaTest=predict(modFitLda, newdata=testing2)
predLdaTest
```

## Reference
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H.
Qualitative Activity Recognition of Weight Lifting Exercises.
Proceedings of 4th International Conference in Cooperation with SIGCHI
(Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.



